<!-- Each chapter is set to compile separately - include "global" set-up -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if ( !require("tidyverse") )   { install.packages("tidyverse") };    require(tidyverse)
if ( !require("kableExtra") )  { install.packages("kableExtra") };   require(kableExtra)
if ( !require("formattable") ) { install.packages("formattable") };  require(formattable)

options(kableExtra.html.bsTable = T)
options(kableExtra.auto_format = FALSE)
```

``` {r Set-Global-Chapter-Variables, echo = FALSE}
plot_bg <- "#FFFFFF" # "#191919"
plot_fg <- "#000000" # "#929292"
plot_fg_alt <- "#969696" # ??
```

# Options Before Expiry {#Before-Expiry}

```{r Current-Price, echo = FALSE}
Apple_Current_Price <- 246.58
```

Our next step is to look at payoff diagrams of the options (strategies) that we've considered before their expiry date, to build intuition on how they might behave. We'll be using this intuition to help us understand the option greeks, the main risk management tool of an options trader, in the next chapter. As always, we'll use real market data to build our plots. We'll stick with Apple, expect this time the option prices are as of the market close on October 25, 2019. Apple is trading at `r currency(Apple_Current_Price)`, slightly higher than in previous chapters. We'll use similar strike prices as before, and consider options that expire on December 20, 2019, March 20, 2020, September 18, 2020, and January 15, 2021. We use the midprice of the option, except when the ask price is substantially higher than the ask prices of surrounding strikes. When that happens, we use the ask price of the option with next closest strike (higher for calls, and lower for puts). 

```{r Different-Strikes-and-Expiries, echo = FALSE}
Apple_Calls_K <- c(180, 210, 240, 270, 300, 330)
Apple_Puts_K <- c(180, 210, 240, 270, 300, 330)

Apple_Current_Price <- 246.58
Apple_Future_Prices <- c(150, 200, 225, 250, 275, 300, 350) 

# Call Options 
Apple_Calls_Cost_Dec19inOct19 <- c( (65.25+68.85)/2, (35.40+39.30)/2, (12.25+12.80)/2,
                                    (1.34+1.82)/2, (0.13+0.14)/2, (0.00+0.02)/2 )
# Apple_Calls_Cost_Dec19inDec19

Apple_Calls_Cost_March20inOct19 <- c( (65.60+69.50)/2, (39.30+42.40)/2, (18.35+19.65)/2,
                                      (5.50+5.80)/2, (1.25+1.64)/2, NA/2 )
# Apple_Calls_Cost_March20inDec19

Apple_Calls_Cost_Sept20inOct19 <- c( (68.35+72.75)/2, (44.10+48.45)/2, (26.50+27.70)/2,
                                     (13.00+14.75)/2, (5.50+6.00)/2, (2.15+2.47)/2 )
# Apple_Calls_Cost_Sept20inDec19

Apple_Calls_Cost_Jan21inOct19 <- c( (72.30+73.10)/2, (49.35+50.15)/2, (30.85+31.60)/2,
                                    (17.50+18.00)/2, (8.85+9.55)/2, (4.25+4.75)/2 )
# Apple_Calls_Cost_Jan21inDec19

# Put Options 
Apple_Puts_Cost_Dec19inOct19 <- c( (0.24+0.33)/2, (1.09+1.14)/2, (6.05+6.40)/2, 
                                   (23.80+27.30)/2, (51.95+56.30)/2, (81.85+86.20)/2 )
# Apple_Puts_Cost_Dec19inDec19

Apple_Puts_Cost_March20inOct19 <- c( (1.00+1.70)/2, (2.80+3.20)/2, (11.65+12.70)/2,
                                     (27.85+30.35)/2, (52.55+56.70)/2, NA/2 ) 
# Apple_Puts_Cost_March20inDec19

Apple_Puts_Cost_Sept20inOct19 <- c( (4.00+4.25)/2, (7.00+9.65)/2, (19.25+20.65)/2,
                                    (35.50+37.80)/2, (56.90+60.25)/2, (82.50+87.30)/2 )
# Apple_Puts_Cost_March20inDec19

Apple_Puts_Cost_Jan21inOct19 <- c( (5.95+7.45)/2, (10.10+13.30)/2, (23.10+25.60)/2,
                                   (37.50+41.95)/2, (59.20+61.45)/2, (83.20+88.00)/2 )
# Apple_Puts_Cost_March20inDec19
```

## Calls and Puts


## Covered Calls


## Protective Puts
