<!-- Each chapter is set to compile separately - include "global" set-up -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

if ( !require("tidyverse") )     { install.packages("tidyverse") };       require(tidyverse)
if ( !require("htmlwidgets") )   { install.packages("htmlwidgets") };     require(htmlwidgets)
if ( !require("plotly") )        { install.packages("plotly") };          require(plotly)
if ( !require("formattable") )   { install.packages("formattable") };     require(formattable)
if ( !require("kableExtra") )    { install.packages("kableExtra") };      require(kableExtra)

options(kableExtra.html.bsTable = T)
options(kableExtra.auto_format = FALSE)
```

```{r Set-Global-Chapter-Variables, echo = FALSE}
plot_bg <- "#FFFFFF" # "#191919"
plot_fg <- "#000000" # "#929292"
plot_fg_alt <- "#969696" # ??

options_theme <- function() {
  theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5, face="bold", 
                                  margin = margin(10, 0, 10, 0),
                                  colour = plot_fg),
        legend.justification=c(0,1), 
        legend.position=c(0,1), 
        legend.title = element_blank(),
        legend.key = element_rect(fill = plot_bg, color = plot_bg),
        legend.background = element_rect(fill = plot_bg, color = plot_bg),
        legend.box.background = element_rect(fill = plot_bg, color = plot_bg),
        legend.text = element_text(colour = plot_fg),
        text = element_text(size=16), 
        panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.background = element_rect(fill = plot_bg, color = plot_bg),
        plot.background = element_rect(fill = plot_bg, color = plot_bg),
        axis.line = element_line(colour = plot_bg),
        axis.title.x = element_text(margin = unit(c(4, 0, 0, 0), "mm"), 
                                    colour = plot_fg),
        axis.title.y = element_text(margin = unit(c(0, 4, 0, 0), "mm"), 
                                    angle = 90, colour = plot_fg),
        axis.text.x = element_text(colour = "black"), 
        axis.text.y = element_text(colour = "black") )
}
```

# A First Look at Options Strategies {#First-Look}

Now that we're armed with payoff diagrams, we're going to work through understanding some options strategies -- the covered call (owning 100 shares and selling 1 call), the protective put (owning 100 shares and buying 1 put), and the straddle (buying 1 put and 1 call). Just like in Chapter \@ref(Payoff-Diagrams), we'll only be looking at payoff diagrams on the expiry date of the options. This will let us focus on the reasons why an investor would use these strategies, and their advantages and disadvantages compared to owning shares. In subsequent chapters, we'll discuss how to draw payoff diagrams for options before their expiry date, and how to use them to manage our options trades. 



## The Covered Call

```{r Apple-Covered-Call, echo = FALSE}
Apple_Current_Price <- 222.76
Apple_Call_Cost <- 8.50
Apple_Call_K <- 230
```

There are two components to a covered call. The first is that we sell a call. Since the buyer has the option to exercise the call, selling a call obligates the seller to provide the shares at the strike price if the buyer exercises. The second is that we already own the shares that we may need to provide the buyer. This protects us from having to scramble for cash to cover our obligation to the buyer. We'll use the `r currency(Apple_Call_K)` strike call on Apple that expire on December 20, 2019 as our example of a covered call.

The `r currency(Apple_Call_K)` strike call costs `r currency(Apple_Call_Cost)`, so we can sell the `r currency(Apple_Call_K)` call that expires on December 20, 2019 for `r currency(Apple_Call_Cost)`. The payoff diagram (on  December 20, 2019) for someone who has sold this call is:  

```{r Apple-Sell-Call-Option-Payoff, out.width = "80%", fig.align = "center", message = FALSE}
Apple_Call_Cost <- 8.50
Apple_Call_K <- 230
x_min <- 210; x_max = 260

Call_x <- seq(from = x_min, to = x_max, by = 1)
Call_y <- Apple_Call_Cost - (Call_x > Apple_Call_K) * (Call_x - Apple_Call_K)
Call_Payoff_Plot <- tibble( "Call_x" = currency(seq(from = x_min, to = x_max, by = 1)),
                            "Call_y" = currency(Apple_Call_Cost - 
                                                  (Call_x > Apple_Call_K) * 
                                                  (Call_x - Apple_Call_K)) )

ggplot() + 
  geom_line(data = Call_Payoff_Plot, aes(Call_x, Call_y, color = "Shares"), linetype = "solid") +
  geom_line(data = Call_Payoff_Plot, aes(Call_x, rep(0,length(Call_x)), color = "Breakeven"), linetype = "dashed") +
  geom_line(data = Call_Payoff_Plot, aes(x = Apple_Call_K + Apple_Call_Cost, 
                                         y = seq(from = -2.5 * Apple_Call_Cost, 
                                                 to = x_max - Apple_Call_K, length = length(Call_x)), 
                                         color = "Breakeven"),
                                         linetype = "dashed") +
  scale_y_continuous(label = scales::dollar) +
  scale_x_continuous(label = scales::dollar) + 
  labs(title = "Selling Apple's $230 Call",
       x = "Stock Price",
       y = "Profit Per Share") +
  scale_color_manual(values = c('Shares'=plot_fg,'Breakeven'=plot_fg,'Breakeven'=plot_fg),
                     labels = c("Profit Per Share",
                                paste0("Breakeven Price = ", currency(Apple_Call_K + Apple_Call_Cost)),
                                paste0("Breakeven Price = ", currency(Apple_Call_K + Apple_Call_Cost))),
                     guide = "legend") + 
  guides(color = guide_legend(override.aes = list(linetype = c("solid", "dashed")))) + 
  options_theme()
```

Before making our first stab at understanding this payoff diagram, remember that we're on the other side of the trade as the buyer. If we sell this call, we collect the `r currency(Apple_Call_Cost)` the buyer pays for it. If Apple's shares are below `r currency(Apple_Call_K)` on December 20, 2019, the call expires worthless and we keep the `r currency(Apple_Call_Cost)`. This is our best case scenario. If Apple's shares are above `r currency(Apple_Call_K)`, the buyer gets to purchase 100 shares for `r currency(Apple_Call_K)` each. But, remember that the buyer isn't profitable until the shares go above his breakeven of `r currency(Apple_Call_K)` + `r currency(Apple_Call_Cost)` = `r currency(Apple_Call_K+Apple_Call_Cost)`. Flipping that around, the seller will not lose money until Apple's shares are above `r currency(Apple_Call_K+Apple_Call_Cost)`. They will, however, gradually lose the `r currency(Apple_Call_Cost)` of potential profit they could have made by selling the call as Apple's shares move from `r currency(Apple_Call_K)` to `r currency(Apple_Call_K+Apple_Call_Cost)`. 

We're only considering the payoff of the call on its expiry date in this section, but options (on stocks) be excerised at any time. If they are exercised early, the buyer captures the intrinsic value of the call but forfeits the remaining time value. While it doesn't make much sense to forfeit the time value, the buyer can choose to exercise early. If they do, the seller must immediately provide them with 100 shares. If they do not own the shares, the seller must buy them. The lowest price a buyer could exercise the call at is `r currency(Apple_Call_K)`, which would cost the seller `r currency(Apple_Call_K)` x 100 = `r currency(100*Apple_Call_K)`. If you compare `r currency(100*Apple_Call_K)` to the `r currency(Apple_Call_Cost)` x 100 = `r currency(100*Apple_Call_Cost)` the seller collects, you'd see that we'd have to come up with `r currency(100*Apple_Call_K)` - `r currency(100*Apple_Call_Cost)` = `r currency(100*(Apple_Call_K-Apple_Call_Cost))` in cash to buy the rest of the shares we owe the buyer. That's a crazy amount of cash to give up when we can only make `r currency(100*Apple_Call_Cost)`, so you might be thinking: why would anyone sell calls?

Luckily, we can fix this problem by selling covered calls. A covered call is when we sell 1 call but already own 100 shares of the stock. If we have the 100 shares that we may have to give the buyer, we'll never be forced to frantically scramble for cash. Let's look at a payoff diagram of a covered call - buying 100 shares of Apple at `r currency(Apple_Current_Price)` each and selling 1 call for `r currency(Apple_Call_Cost)` x 100 = `r currency(100*Apple_Call_Cost)`.

```{r Apple-Covered-Call-Payoff-Diagram, out.width = "80%", fig.align = "center", message = FALSE}
Apple_Call_Cost <- 8.50
Apple_Call_K <- 230
x_min <- 190; x_max = 260 
Call_x <- seq(from = x_min, to = x_max, by = 1)

Call_Payoff_Plot <- tibble( "Call_x" = currency(seq(from = x_min, to = x_max, by = 1)),
                            "Covered_Call_y" = currency(Apple_Call_Cost + (Call_x - Apple_Current_Price) - 
                                                  pmax(0, Call_x - Apple_Call_K)) ) 

ggplot() + 
  geom_line(data = Call_Payoff_Plot, aes(Call_x, Covered_Call_y, color = "Shares"), 
            linetype = "solid") +
  geom_line(data = Call_Payoff_Plot, aes(Call_x, rep(0,length(Call_x)), color = "Breakeven"), 
            linetype = "dashed") +
  geom_line(data = Call_Payoff_Plot, aes(x = Apple_Current_Price - Apple_Call_Cost, 
                                         y = seq(from = -4*Apple_Call_Cost, to = x_max-Apple_Call_K, 
                                                 length=length(Call_x)), 
                                         color = "Breakeven"),
                                         linetype = "dashed") +
  geom_line(data = Call_Payoff_Plot, aes(Call_x, Call_x - Apple_Current_Price, color = "Own_Shares"), 
            linetype = "solid") +
  scale_y_continuous(label = scales::dollar) +
  scale_x_continuous(label = scales::dollar) + 
  labs(title = "Selling Apple's $230 Covered Call",
       x = "Stock Price",
       y = "Profit Per Share") +
  scale_color_manual(values = c('Shares'=plot_fg,'Breakeven'=plot_fg,
                                'Breakeven'=plot_fg,"Own_Shares"=plot_fg_alt),
                     labels = c("Covered Call - Profit Per Share",
                                "Owning Shares - Profit Per Share",
                                paste0("Breakeven Price = ", currency(Apple_Current_Price - Apple_Call_Cost))),
                     guide = "legend") + 
  guides(color = guide_legend(override.aes = list(linetype = c("solid", "solid", "dashed")))) + 
  options_theme()
```

By selling a call with a strike price of `r currency(Apple_Call_K)`, the call buyer gets all of the gains when Apple's shares are above `r currency(Apple_Call_K)`. This caps our upside of owning the shares at `r currency(Apple_Call_K)`, which is why the per-share profit of a covered call flattens out when the shares reach `r currency(Apple_Call_K)`. We always collect the `r currency(Apple_Call_Cost)` per share we make by selling the call, which is is cash that goes into our account. This is why the covered call (black line) makes `r currency(Apple_Call_Cost)` more profit than owning shares (orange line) whenever Apple's shares are below `r currency(Apple_Call_K)`. What's even better is that the `r currency(Apple_Call_Cost)` also reduces our breakeven to `r currency(Apple_Current_Price)` - `r currency(Apple_Call_Cost)` = `r currency(Apple_Current_Price-Apple_Call_Cost)`. 

Another way to think of this payoff diagram is that the covered call has two sources of return - the money collected by selling the call, and the remaining upside participation we have from owning shares. In this case, we collect `r currency(Apple_Call_Cost)` by selling the call and maintain `r currency(Apple_Call_K)` - `r currency(Apple_Call_Cost)` = `r currency(Apple_Call_K-Apple_Call_Cost)` of upside participation in the shares. We have a total of `r currency(Apple_Call_Cost)` + `r currency(Apple_Call_K-Apple_Current_Price)` = `r currency(Apple_Call_Cost+Apple_Call_K-Apple_Current_Price)` in upside. If Apple's shares are above `r currency(Apple_Current_Price)` + `r currency(Apple_Call_Cost+Apple_Call_K-Apple_Current_Price)` = `r currency(Apple_Current_Price+Apple_Call_Cost+Apple_Call_K-Apple_Current_Price)`, we would make more money if we had not sold the calls. This value is marked in the payoff diagram by where the orange and black lines cross. 

Putting these together, we lose less money if Apple's shares fall below `r currency(Apple_Current_Price-Apple_Call_Cost)` and make less money (compared to if we had not sold the call) if they go above `r currency(Apple_Current_Price+Apple_Call_Cost)` That means the covered call strategy reduces the *volatility* of our investment, which can be very desirable. It also produces income - the `r currency(Apple_Call_Cost)` is cash that goes straight into our account. We bought the shares for `r currency(Apple_Current_Price)`, which means we made a 3-month yield of <span class = "fraction" data-top="`r currency(Apple_Call_Cost)`" data-bottom="`r currency(Apple_Current_Price)`"></span> = `r percent(Apple_Call_Cost/Apple_Current_Price)` (while Treasuries barely yield under 2\% per year).

This ability to generate income is one of the covered call's most appealing attributes. For investors who seek to generate consistent (inflation-adjusted) income from their portfolio each year, the volatility-reducing characteristics of the covered call are also welcome. Let's go back to our trusty table, and compare owning 100 shares of Apple to owning 100 shares and selling 1 call. As before, the shares are bought at `r currency(Apple_Current_Price)`. We'll consider the profit or loss when the call is sold at a variety of different strike prices. For context, remember that: 

```{r Apple-Options-Cost}
Apple_Calls_K <- c(145, 180, 205, 230)
Apple_Puts_K <- c(145, 180, 205, 230)
Apple_Calls_Cost <- c(78.45, 44.85, 23.63, 8.50)
Apple_Puts_Cost <- c(0.32, 1.69, 5.45, 15.30)
```

1. 1 `r currency(Apple_Calls_K[1])` strike call costs `r currency(Apple_Calls_Cost[1])` x 100 = `r 100*currency(Apple_Calls_Cost[1])`.
2. 1 `r currency(Apple_Calls_K[2])` strike call costs `r currency(Apple_Calls_Cost[2])` x 100 = `r 100*currency(Apple_Calls_Cost[2])`.
3. 1 1 `r currency(Apple_Calls_K[3])` strike call costs `r currency(Apple_Calls_Cost[3])` x 100 = `r 100*currency(Apple_Calls_Cost[3])`.
4. 1 1 `r currency(Apple_Calls_K[4])` strike call costs `r currency(Apple_Calls_Cost[4])` x 100 = `r 100*currency(Apple_Calls_Cost[4])`.

```{r Apple-Covered-Call-Different-Strikes, out.width = "80%", fig.align = "center", message = FALSE}
Apple_Future_Prices <- c(100, 150, 200, 220, 245, 300)
Apple_Current_Price <- 222.76

Apple_Covered_Call <- tibble( "Share Price" = scales::dollar(Apple_Future_Prices, negative_parens = TRUE),
      "Buy 100 Shares" = scales::dollar( 100 * ( Apple_Future_Prices - Apple_Current_Price ),
                                         negative_parens = TRUE ), 
      "Buy 100 Shares, Sell 1 $145 Strike Call" = scales::dollar( 
        100 * ( ( Apple_Calls_Cost[1] ) + ( Apple_Future_Prices - Apple_Current_Price ) -
                  pmax( 0, Apple_Future_Prices - Apple_Calls_K[1] ) ),
        negative_parens = TRUE ),
      "Buy 100 Shares, Sell 1 $180 Strike Call" = scales::dollar( 
        100 * ( ( Apple_Calls_Cost[2] ) + ( Apple_Future_Prices - Apple_Current_Price ) -
                  pmax( 0, Apple_Future_Prices - Apple_Calls_K[2] ) ),
        negative_parens = TRUE ),
      "Buy 100 Shares, Sell 1 $205 Strike Call" = scales::dollar( 
        100 * ( ( Apple_Calls_Cost[3] ) + ( Apple_Future_Prices - Apple_Current_Price ) -
                  pmax( 0, Apple_Future_Prices - Apple_Calls_K[3] ) ),
        negative_parens = TRUE ),
      "Buy 100 Shares, Sell 1 $230 Strike Call" = 
        scales::dollar( 100 * ( ( Apple_Calls_Cost[4] ) + ( Apple_Future_Prices - Apple_Current_Price ) -
                                  pmax( 0, Apple_Future_Prices - Apple_Calls_K[4] ) ),
                        negative_parens = TRUE ) )

kable_styling( knitr::kable(Apple_Covered_Call, 
                            caption = "Dollar Return if Held to Expiry"), 
               bootstrap_options = c("bordered", "condensed", "responsive"),
               full_width = FALSE )
```

The first thing to notice is that the gains (or losses) of every covered call position is lower than just owning shares. Covered call compress the range of possible gains or losses, which reduces our *volatility*. Next, we'll examine the risk-reward tradeoff of selling covered calls at different strike prices. As always, the best way to do this is by drawing a diagram. I'm only showing the covered call with a `r currency(Apple_Calls_K[1])` strike. I'd strongly encourage drawing the the remaining diagrams, so that the next few paragraphs don't turn into an incomprehensible blur of numbers.

```{r Apple-145-Covered-Call-Payoff-Diagram, out.width = "80%", fig.align = "center", message = FALSE}
Apple_Call_Cost <- 78.45
Apple_Call_K <- 145
Apple_Current_Price <- 222.76

x_min = 120; x_max = 165
Call_x <- seq(from = x_min, to = x_max, by = 1)
Call_Payoff_Plot <- tibble( "Call_x" = currency(seq(from = x_min, to = x_max, by = 1)),
        "Covered_Call_y" = currency(Apple_Call_Cost + (Call_x - Apple_Current_Price) - 
                                      pmax(0, Call_x - Apple_Call_K)) ) 

ggplot() + 
  geom_line(data = Call_Payoff_Plot, aes(Call_x, Covered_Call_y, color = "Shares"),
            linetype = "solid") +
  geom_line(data = Call_Payoff_Plot, aes(Call_x, rep(0,length(Call_x)), color = "Breakeven"), 
            linetype = "dashed") +
  geom_line(data = Call_Payoff_Plot, aes(x = Apple_Current_Price - Apple_Call_Cost, 
                                         y = seq(from = -1.25*Apple_Call_Cost, to = 1.25*x_max-Apple_Call_K, 
                                                 length=length(Call_x)), 
                                         color = "Breakeven"),
                                         linetype = "dashed") +
  geom_line(data = Call_Payoff_Plot, aes(Call_x, Call_x - Apple_Current_Price, color = "Own_Shares"), 
            linetype = "solid") +
  scale_y_continuous(label = scales::dollar) +
  scale_x_continuous(label = scales::dollar) + 
  labs(title = "Selling Apple's $145 Covered Call",
       x = "Stock Price",
       y = "Profit Per Share") +
  scale_color_manual(values = c('Shares'=plot_fg,'Breakeven'=plot_fg,
                                'Breakeven'=plot_fg,"Own_Shares"=plot_fg_alt),
                     labels = c("Covered Call - Profit Per Share",
                                "Owning Shares - Profit Per Share",
                                paste0("Breakeven Price = ", 
                                       currency(Apple_Current_Price - Apple_Call_Cost))),
                     guide = "legend") + 
  guides(color = guide_legend(override.aes = list(linetype = c("solid", "solid", "dashed")))) + 
  options_theme()
```

This payoff diagram looks strange for two reasons. The first is that the black line barely gets above `r currency(0)` we make so little money after shares go above the breakeven price, and the second is that we lose so much more money if we just owned shares. This puts the grey

The `r currency(Apple_Call_K)` strike covered call yields `r currency(Apple_Call_Cost)` x 100 = `r 100*currency(Apple_Call_Cost)`. If the shares fall to `r currency(Apple_Future_Prices[1])` on December 20, 2019, we lose `r currency(Apple_Future_Prices[1])` - `r currency(Apple_Current_Price)` = `r currency(Apple_Future_Prices[1]-Apple_Current_Price)` per share, or 100 x `r currency(Apple_Future_Prices[1]-Apple_Current_Price)` = `r currency(Apple_Future_Prices[1]-Apple_Current_Price)` from owning 100 shares. Adding the `r 100*currency(Apple_Call_Cost)` we collected from selling the call and the `r 100*currency(Apple_Future_Prices[1]-Apple_Current_Price)` we lost from owning the shares, we lose a total of `r 100*currency(Apple_Call_Cost)` - `r -100*currency(Apple_Future_Prices[1]-Apple_Current_Price)` = `r 100*currency(Apple_Call_Cost+(Apple_Future_Prices[1]-Apple_Current_Price))`, which matches the number in the table. 

The breakeven of the `r currency(Apple_Call_K)` covered call is `r currency(Apple_Current_Price)` - `r currency(Apple_Call_Cost)` = `r currency(Apple_Current_Price-Apple_Call_Cost)`, and we receive a flat payoff of \$69 when Apple's shares are above breakeven. If we think Apple's shares are going to be above `r currency(Apple_Current_Price-Apple_Call_Cost)` on December 20, 2019, we are generating `r currency( 100 * ( ( Apple_Calls_Cost[1] ) + ( Apple_Future_Prices - Apple_Current_Price ) - pmax( 0, Apple_Future_Prices - Apple_Calls_K[1] ) )[2] )` of income from a cash investment of `r currency(100 * Apple_Current_Price)` That's a pitifully low 3-month yield of <span class = "fraction" data-top="`r 100 * currency( ( ( Apple_Calls_Cost[1] ) + ( Apple_Future_Prices - Apple_Current_Price ) - pmax( 0, Apple_Future_Prices - Apple_Calls_K[1] ) )[2] )`" data-bottom="`r 100*currency(Apple_Current_Price)`"></span> = `r percent( ( ( Apple_Calls_Cost[1] ) + ( Apple_Future_Prices - Apple_Current_Price ) - pmax( 0, Apple_Future_Prices - Apple_Calls_K[1] ) )[2] / Apple_Current_Price)`. So while the `r currency(Apple_Call_K)` covered call has an extremely low breakeven of, it has a 3-month yield of only `r percent( ( ( Apple_Calls_Cost[1] ) + ( Apple_Future_Prices - Apple_Current_Price ) - pmax( 0, Apple_Future_Prices - Apple_Calls_K[1] ) )[2] / Apple_Current_Price)`. So while the `r currency(Apple_Call_K)`. If we're selling covered calls, we probably want more income than that. Moving the strike of the call higher lets us get more income, but at the cost of having a higher breakeven price. 

Let's look at a `r currency(Apple_Calls_K[2])` strike covered call. It has a breakeven of `r currency(Apple_Current_Price)` - `r currency(Apple_Calls_Cost[2])` = `r currency(Apple_Current_Price-Apple_Calls_Cost[2])`, which is `r currency(Apple_Current_Price-Apple_Calls_Cost[2])` - `r currency(Apple_Current_Price-Apple_Call_Cost)` = `r currency( (Apple_Current_Price-Apple_Calls_Cost[2]) - (Apple_Current_Price-Apple_Call_Cost) )` more than the `r currency(Apple_Call_K)` covered call. This is because it gives `r currency(Apple_Calls_Cost[1])` - `r currency(Apple_Calls_Cost[2])` = `r currency(Apple_Calls_Cost[1]-Apple_Calls_Cost[2])` less income per share than the `r currency(Apple_Call_K)` covered call. Since the `r currency(Apple_Current_Price-Apple_Calls_Cost[2])` covered call gives less income, it is less able to buffer losses that can arise from owning shares. For example, when Apple's shares are at `r currency(Apple_Future_Prices[1])`, the `r currency(Apple_Calls_K[2])` covered call loses `r 100*currency(Apple_Call_Cost)` and the `r currency(Apple_Call_K)` covered call loses `r -100*currency(Apple_Call_Cost+(Apple_Future_Prices[1]-Apple_Current_Price))`. In exchange for having more downside risk than the `r currency(Apple_Calls_K[1])` covered call, the `r currency(Apple_Calls_K[2])` covered call offers a higher 3-month yield of <span class = "fraction" data-top="`r currency( 100 * ( ( Apple_Calls_Cost[2] ) + ( Apple_Future_Prices - Apple_Current_Price ) - pmax( 0, Apple_Future_Prices - Apple_Calls_K[2] ) )[3] )`" data-bottom="`r 100*currency(Apple_Current_Price)`"></span> = `r percent( ( ( Apple_Calls_Cost[2] ) + ( Apple_Future_Prices - Apple_Current_Price ) - pmax( 0, Apple_Future_Prices - Apple_Calls_K[2] ) )[3] / Apple_Current_Price)`.

That is more than we can make in a savings account or Treasury bond, and our breakeven is reasonably low. Apple would have to fall by slightly over `r percent(20/100)` in 3 months to fall below breakeven. While Apple could fall by `r percent(20/100)`, it's pretty unlikely. We have already considered the payoff diagram of the `r currency(Apple_Calls_K[4])` covered call - it had a breakeven of `r currency(Apple_Current_Price - Apple_Calls_Cost[4])` and can reasonably be expected to yield over `r percent(10/100)` per year. This is much higher than the `r percent( ( ( Apple_Calls_Cost[1] ) + ( Apple_Future_Prices - Apple_Current_Price ) - pmax( 0, Apple_Future_Prices - Apple_Calls_K[1] ) )[2] / Apple_Current_Price)` yield of the `r currency(Apple_Calls_K[1])` covered call or the `r percent( ( ( Apple_Calls_Cost[2] ) + ( Apple_Future_Prices - Apple_Current_Price ) - pmax( 0, Apple_Future_Prices - Apple_Calls_K[2] ) )[3] / Apple_Current_Price)` yield of the `r currency(Apple_Calls_K[2])` covered call, it comes at the cost of having a much higher breakeven price. The key takeaway is that we can achieve covered calls at the  higher yields  more yield comes at the cost of having a higher breakeven prices. 

```{r Apple-Covered-Call-Different-Strikes-Relative, out.width = "80%", fig.align = "center", message = FALSE}
Apple_Future_Prices <- c(100, 150, 200, 220, 245, 300)
Apple_Current_Price <- 222.76
Apple_Future_Returns <-  Apple_Future_Prices - Apple_Current_Price

Apple_Covered_Call_Relative <- tibble( "Share Price" = Apple_Future_Prices,
      "Buy 100 Shares" = sign(Apple_Future_Returns) * 
        percent( ( Apple_Future_Prices - Apple_Current_Price ) / 
                   ( Apple_Future_Prices - Apple_Current_Price ) ), 
      "Buy 100 Shares, Sell 1 $145 Strike Call" = sign(Apple_Future_Returns) * 
        percent( ( ( Apple_Calls_Cost[1] ) + ( Apple_Future_Prices - Apple_Current_Price ) - 
                     pmax( 0, Apple_Future_Prices - Apple_Calls_K[1] ) ) / 
                   ( Apple_Future_Prices - Apple_Current_Price ) ),
      "Buy 100 Shares, Sell 1 $180 Strike Call" = sign(Apple_Future_Returns) * 
        percent( ( ( Apple_Calls_Cost[2] ) + ( Apple_Future_Prices - Apple_Current_Price ) - 
                     pmax( 0, Apple_Future_Prices - Apple_Calls_K[2] ) ) / 
                   ( Apple_Future_Prices - Apple_Current_Price ) ),
      "Buy 100 Shares, Sell 1 $205 Strike Call" = sign(Apple_Future_Returns) * 
        percent( ( ( Apple_Calls_Cost[3] ) + ( Apple_Future_Prices - Apple_Current_Price ) - 
                     pmax( 0, Apple_Future_Prices - Apple_Calls_K[3] ) ) / 
                   ( Apple_Future_Prices - Apple_Current_Price ) ),
      "Buy 100 Shares, Sell 1 $230 Strike Call" = sign(Apple_Future_Returns) * 
        percent( ( ( Apple_Calls_Cost[4] ) + ( Apple_Future_Prices - Apple_Current_Price ) - 
                     pmax( 0, Apple_Future_Prices - Apple_Calls_K[4] ) ) / 
                   ( Apple_Future_Prices - Apple_Current_Price ) ) )

kable_styling( knitr::kable(Apple_Covered_Call_Relative, 
                            caption = "Relative Return if Held to Expiry"), 
               bootstrap_options = c("bordered", "condensed", "responsive"),
               full_width = FALSE )
```

As before, the percentage returns when shares are at `r currency(Apple_Future_Prices[3])` is somewhat misleading, since someone who bought 100 shares of Apple's stock at `r currency(Apple_Current_Price)` has only lost `r 100*currency( Apple_Future_Prices[4] - Apple_Current_Price )`. The relative returns are large and positive when the shares are at `r currency(Apple_Future_Prices[3])` because the income from selling the call is more than the losses incurred from holding shares originally purchased at `r currency(Apple_Current_Price)`. More generally, the table of relative returns shows smaller losses *and* less upside when we sell covered calls with lower strike prices. 

There is one deficiency in this table - the highest strike price it considers is `r currency(Apple_Calls_K[4])`, which is not much higher than the original purchase price of `r currency(Apple_Current_Price)`. While we don't have its precise price, a `r currency(300)` strike call with the same expiry date will cost less than `r currency(1)`. With Apple trading near all-time highs at `r currency(Apple_Current_Price)`, the chance of it reaching `r currency(300)` in 3 months is quite low. So we are very likely to keep the `r currency(1)` we collect by selling the call, for a pitifully low yield 3-month yield of <span class = "fraction" data-top="`r currency(1)`" data-bottom="`r currency(Apple_Current_Price)`"></span> = `r percent( 1 / Apple_Current_Price )`.

Our breakeven price and our yield are related in an even more important way. This is something I like to call re-investment risk - that's where we make less money from an investing strategy over time. For covered calls, this can happen if the shares we own decline over time. We are using the shares we own to protect against the risk of being called, so selling our shares before the call expires defeats the purpose of the strategy. And after the call expires, we will have to sell another covered call to collect more yield. Here's the problem - every time we sell a call, the best short-term outcome is that the call expires worthless so that we keep all the cash that we collected. But if our shares gradually drift downward over time, we will be selling our calls against a smaller total investment. That means that even if the yield stays the same, we'll collect less income over time. 

<!--
Putting some numbers behind those words, let's consider  example of Apple falling roughly `r percent(10)` from it's current price of `r currency(Apple_Current_Price)` to `r 0.90*currency(Apple_Current_Price)` on ** Add date **. The ** Finish example **.
-->

Another idea that we'll spend lots of time with later is how we can use options to adjust our market positioning. Since options are time-limited, the market exposure they give us naturally resets when they expire. Depending on if prices move in our favor by then, this can be good or bad. But options also give us the flexibility to adjust our positioning before they expire. Let's use the covered call as a simple example of this *flexibity*. If Apple is at `r currency(Apple_Current_Price)` and we sell a `r currency(Apple_Calls_K[4])` covered call, we reduce our breakeven from `r currency(Apple_Current_Price)` to `r currency(Apple_Current_Price)` - `r currency(Apple_Calls_Cost[4])` = `r currency(Apple_Current_Price-Apple_Calls_Cost[4])` *if* we hold position until the call expires on December 20, 2019. However, if Apple's stock jumps above our breakeven price of `r currency(Apple_Current_Price+Apple_Calls_Cost[4])` weeks before December 20, the call will increase by much less than the 100 shares do. Since we sold the call and own the shares, we'll be able to close the covered call early and still make money. This is because of the *delta* of the call, which we'll take on in the next chapter. Even though we'll make less money than if we didn't sell the call, this shows there are scenarios where we can still make money with options even when the trade moves against us. 

Stepping back and looking at the big picture for the covered call, the strike price we chose to sell the call at specifies how much risk we're taking on. The key challenge in using the covered call strategy is that we have to balance two competing goals: 

1. Protecting against *re-investment risk* so that we can maximize the long-term performance of the covered call strategy. 
2. Maximizing the yield of every call we sell. 

This section has attempted to demonstrate the income-generating potential of covered calls, their volatility-reducing characteristics, and the nature of the risk-reward tradeoff we make when chosing which strike to sell the covered call at. We've also introduced two concepts -- *re-investment risk* and the *flexibility of options* -- that we'll come back to later on. 


## The Protective Put

A protective put is buying 100 shares and 1 put of the same stock. The protective put changes the risk-reward characteristics of our trade - the put cancels out some of the potential downside of owning shares. There are other ways to limit downside risk, like selling 50 of our 100 shares. If we do that, we'll lose half as much money if Apple's shares go down *and* save the money we would have spent on puts. But we would also lose the potential gains those 50 shares could give - we'd only make half as much money if they go up. If we own all 100 shares, we keep all of the upside. And pairing the 100 shares with 1 put typically limits our downside to just a few percent. 

Let's compare the possible gains and losses of three different investments - 100 shares of Apple, 50 shares of Apple, and 100 shares and 1 `r currency(Apple_Calls_K[3])` strike put of Apple. To help us stay focused on the big picture, we'll just be looking at the price of the put on its expiry date of December 20, 2019. As always, our first shot at understanding this strategy is by looking at its payoff diagram.

```{r Apple-Protective-Put-205-Strike-Payoff-Diagram, out.width = "80%", fig.align = "center", message = FALSE}
Apple_Put_Cost <- 5.45
Apple_Put_K <- 205
Apple_Current_Price <- 222.76

x_min = 180; x_max = 260
Put_x <- seq(from = x_min, to = x_max, by = 1)
Put_Payoff_Plot <- tibble( "Put_x" = currency(seq(from = x_min, to = x_max, by = 1)),
                            "Protective_Put_y" = currency(-Apple_Put_Cost + 
                                                            ( Put_x - Apple_Current_Price ) + 
                                                            pmax(0, Apple_Put_K - Put_x)) ) 

ggplot() + 
  geom_line(data = Put_Payoff_Plot, aes(Put_x, Protective_Put_y, 
                                        color = "Shares"), linetype = "solid") +
  geom_line(data = Put_Payoff_Plot, aes(Put_x, rep(0,length(Put_x)), 
                                        color = "Breakeven"), linetype = "dashed") +
  geom_line(data = Put_Payoff_Plot, aes(x = Apple_Current_Price + Apple_Put_Cost, 
                                         y = seq(from = -8*Apple_Put_Cost, 
                                                 to = 0.8*(x_max-Apple_Put_K), 
                                                 length=length(Put_x)), 
                                         color = "Breakeven"),
                                         linetype = "dashed") +
  geom_line(data = Put_Payoff_Plot, aes(Put_x, Put_x - Apple_Current_Price, color = "Own_Shares"), 
            linetype = "solid") +
  scale_y_continuous(label = scales::dollar) +
  scale_x_continuous(label = scales::dollar) + 
  labs(title = "Buying Apple's $205 Protective Put",
       x = "Stock Price",
       y = "Profit Per Share") +
  scale_color_manual(values = c('Shares'=plot_fg,'Breakeven'=plot_fg,
                                'Breakeven'=plot_fg,"Own_Shares"=plot_fg_alt),
                     labels = c("Protective Put - Profit Per Share",
                                "Owning Shares - Profit Per Share",
                                paste0("Breakeven Price = ", currency(Apple_Current_Price + Apple_Put_Cost)) ),
                     guide = "legend") + 
  guides(color = guide_legend(override.aes = list(linetype = c("solid", "solid", "dashed")))) + 
  options_theme()
```

As a quick reminder, we bought Apple's shares at `r currency(Apple_Current_Price)` on September 11, 2019 and a `r currency(Apple_Puts_K[3])` strike put that expires on December 20, 2019 and costs `r currency(Apple_Puts_Cost[3])`. The breakeven is `r currency(Apple_Current_Price)` + `r currency(Apple_Puts_Cost[3])` = `r currency(Apple_Current_Price+Apple_Puts_Cost[3])`, but this doesn't tell the whole story of the protective put. Whenever Apple's shares are `r currency(Apple_Puts_Cost[3])` or below, we can only lose `r currency(Apple_Puts_K[3])` - `r currency(Apple_Current_Price)` = `r currency(Apple_Puts_K[3]-Apple_Current_Price)` from owning the shares plus `r currency(Apple_Puts_Cost[3])` from purchasing the put, for a maximum total loss of `r currency(Apple_Puts_K[3])` - `r currency(Apple_Current_Price)` - `r currency(Apple_Puts_Cost[3])` = `r currency(Apple_Puts_K[3]-Apple_Current_Price-Apple_Puts_Cost[3])`. This is the *magical* property of the put - on it's expiry date, a put goes up by a dollar for every dollar Apple's stock goes down. We bought a put with a strike price of `r currency(Apple_Puts_K[3])`. Once Apple's stock is below `r currency(Apple_Puts_K[3])`, the put cancels out, dollar-for-dollar, any loses from owning Apple's shares. 

This is captured by the flat portion of the black line in the payoff diagram, which shows that the protective put has a maximum loss of `r -currency(Apple_Puts_K[3]-Apple_Current_Price-Apple_Puts_Cost[3])` per share. By buying the put at `r currency(Apple_Puts_K[3])`, we are effectively removing any downside from Apple's shares below `r currency(Apple_Puts_K[3])` (if we hold the put to expirt). If we just own shares, we'll still lose money if Apple's shares are below `r currency(Apple_Puts_K[3])` (on the expiry date). This is shown by the orange line in the payoff diagram, which just keeps going down. This protection isn't free - it cost `r currency(Apple_Puts_Cost[3])`, or <span class = "fraction" data-top="`r currency(Apple_Puts_Cost[3])`"  data-bottom="`r currency(Apple_Current_Price)`"></span> = `r percent( Apple_Puts_Cost[3] / Apple_Current_Price )` of our original investment. 

This looks super expensive - the put expires in 3 months, so we'll have to buy 4 puts per year to maintain our protection. If they all cost around `r percent(2.5/100)` of our investment, we'll be spending almost `r percent(10/100)` of the investment each year on puts. Even if we think it's expensive, we might still buy the put if we think Apple's shares are more likely to go down over the next 3 months. There might be special circumstances, like Apple missing earnings or a possible recession, that make us think the put is worth its cost. But holding the shares and waiting for them to go back up will probably work just as well, and it's free. So even if we thought Apple's shares were going below `r currency(Apple_Put_K)` before the put expires, why should we buy the put? 

There are two main reasons. The first is practical. Stocks tend to go up over the long run, and trying to make money betting that stocks will go down is probably not a good idea. But stocks are volatile. They can go down for a few months or even years, depending on how the economy is doing and how competitive their industry is. We'd like to be able to stay invested for the long run but sudden expenses can pop up and derail our plans. Or maybe we see an investment that we like more, and would like to own shares of this other company instead. If Apple is down to `r currency(Apple_Puts_K[2])`, which it has reached a few times in the last year, we'd be down <span class = "fraction" data-top="`r currency(Apple_Current_Price)` - `r currency(Apple_Puts_K[2])`"  data-bottom="`r currency(Apple_Current_Price)`"></span> = `r percent( ( Apple_Current_Price - Apple_Puts_K[2] ) / Apple_Current_Price )`, which is almost `r percent(20/100)`. Nobody wants to sell when they're down `r percent(20/100)`, whether it's an actual loss of `r percent(20/100)` or a `r percent(20/100)` loss from the highest price Apple has reached while they've owned it.

Buying the put lets us avoid this situation. When we chose the strike price of the put, we're defining the maximum loss that we're willing to take during those 3 month. This gives us much more *flexibility* than if we have to wait for months for Apple's shares to go back up. If we value this *flexibility*, then buying the put may be reasonable. There are lots of reasons why we might want to preserve flexibility. For example, we could be looking at buying a home in the near future, having to pay for college, be concerned about the health of a family member, or be concerned about a potential layoff. Depending on how much cash we think we might need to help us get through these situations, buying a put will help us preserve the financial flexibility we need to navigate them. 

The second reason is psychological. While stocks tend to go up other the long run, they are volatile. And seeing your account down `r percent(20/100)` is very different than just knowing it can happen, especially if we're bombarded by bad news about our investments or the economy while this happens. It's hard to make good decisions when you're losing money. Even if we manage to do it, we probably doesn't need the extra stress. Buying a put lets us define the maximum loss that we're comfortable with, so it lets us stay in our positions stress-free. While buying protective puts reduces our return (by the cost of the put), reducing our stress will help us make good decisions if the stock market starts dropping. In my opinion, this is the biggest benefit of protective puts.

And remember, we don't have to buy puts all the time. Just when we think the benefits they provide are worth their costs.

```{r Apple-Protective-Put-205-Strike-Table, out.width = "80%", fig.align = "center", message = FALSE}
Apple_Future_Prices <- c(100, 150, 200, 220, 245, 300)
Apple_Current_Price <- 222.76
Apple_Put_K <- 205

Apple_Protective_Put <- tibble( "Share Price" = scales::dollar(Apple_Future_Prices, negative_parens = TRUE),
     "Buy 100 Shares" = scales::dollar(100*(Apple_Future_Prices - Apple_Current_Price), 
                                       negative_parens = TRUE), 
     "Buy 50 Shares" = scales::dollar(50*(Apple_Future_Prices - Apple_Current_Price), 
                                      negative_parens = TRUE),
     "Buy 100 Shares, Buy 1 $205 Strike Put" = scales::dollar( 
       100 * (-Apple_Puts_Cost[3] + (Apple_Future_Prices - Apple_Current_Price) + 
                pmax(0, Apple_Put_K - Apple_Future_Prices)), 
       negative_parens = TRUE) )

kable_styling (knitr::kable(Apple_Protective_Put, escape = FALSE,
                            caption = "Dollar Return if Held to Expiry"),
               bootstrap_options = c("bordered", "condensed", "responsive"),
               full_width = FALSE )
```

This table emphasizes the idea that using protective puts lets us define the maximum loss we can take on an investment, while maintaining almost all of our upside participation. This protective put defines our maximum loss at `r -100*currency(Apple_Puts_K[3]-Apple_Current_Price-Apple_Puts_Cost[3])`. If we own 100 shares of Apple, we'd be down that much when the shares fall by about `r percent(10/100)` to `r currency(200)`. If we own 50 shares, shares have to fall by about `r percent(20/100)` to `r currency(180)`. But if we cut our shares from 100 to 50, we make half as much money when Apple's shares are at `r currency(Apple_Future_Prices[5])` or `r currency(Apple_Future_Prices[6])` on December 20, 2019. This shows that if we want to keep more upside, selling shares is not a good option. 

The breakeven of the protective put is `r currency(Apple_Current_Price)` + `r currency(Apple_Puts_Cost[3])` = `r currency(Apple_Current_Price+Apple_Puts_Cost[3])`. Compared to owning 50 shares, the protective put keeps more upside when shares are at much higher levels like `r currency(Apple_Future_Prices[5])` or `r currency(Apple_Future_Prices[6])`. And once the shares are below about `r currency( (Apple_Future_Prices[2]+Apple_Future_Prices[3])/2 )`, it has less downside. Compared to owning 100 shares, the protective put keeps almost all of the upside. We only reduce our potential gains by the cost of the put, and gain a well-defined maximum loss for our troubles. As always, the options strategy does poorly if Apple's shares stay roughly the same. We bought that shares at `r currency(Apple_Current_Price)` and lose more money (on the expiry date) when the shares are at `r currency(Apple_Future_Prices[4])`. This is because the shares have not gone down enough for the put to make money, and have not gone up enough for our gains to offset the cost of the put. 

How do we chose the strike price of the protective put? By looking at the payoff diagram, the strike price of the put defines what the maximum loss we can take on our investment. With a strike price of `r currency(Apple_Put_K)`, we lose `r -100*currency(Apple_Puts_K[3]-Apple_Current_Price-Apple_Puts_Cost[3])` when the shares are at `r currency(Apple_Put_K)` and `r -100*currency(Apple_Puts_K[3]-Apple_Current_Price-Apple_Puts_Cost[3])` at any price below `r currency(Apple_Put_K)` By thinking through the maximum loss we're willing to tolerate, we can chose the strike price (and expiry date) of the put. 

### A Real Trade - NLY 

```{r NLY-Protective-Put}
NLY_Put_Cost <- 0.66
NLY_Put_K <- 9
NLY_Current_Price <- 8.90
NLY_Div <- 0.25
```

In addition to its other benefits, using protective puts can help us create trades with an attractive risk-reward propostion. For example, on September 24, 2019, Annaly Capital Management (ticker NLY) is trading at `r currency(NLY_Current_Price)` and the January 17, 2020 put with a strike price of `r currency(NLY_Put_K)` is trading for `r currency(NLY_Put_Cost)`. To make this trade even more appealing, NLY is scheduled to make two dividend payments of `r currency(NLY_Div)` each between September 24, 2019 and January 17, 2020. Assuming that NLY makes these two dividend payments, the payoff diagram for this protective put is: 

```{r NLY-Protective-Put-Plot, out.width = "80%", fig.align = "center", message = FALSE}
x_min <- 7; x_max <- 11
Put_x <- seq(from = x_min, to = x_max, by = 0.1)
Put_Payoff_Plot <- tibble( "Put_x" = currency(seq(from = x_min, to = x_max, by = 1)),
                            "Protective_Put_y" = currency(-NLY_Put_Cost + 2*NLY_Div +
                                                            ( Put_x - NLY_Current_Price ) + 
                                                            pmax(0, NLY_Put_K - Put_x)) ) 

ggplot() + 
  geom_line(data = Put_Payoff_Plot, aes(Put_x, Protective_Put_y, 
                                        color = "Shares"), linetype = "solid") +
  geom_line(data = Put_Payoff_Plot, aes(Put_x, rep(0,length(Put_x)), 
                                        color = "Breakeven"), linetype = "dashed") +
  geom_line(data = Put_Payoff_Plot, aes(x = NLY_Current_Price + NLY_Put_Cost - 2*NLY_Div, 
                                         y = seq(from = -4*NLY_Put_Cost, 
                                                 to = 3*(x_max-NLY_Put_K), 
                                                 length=length(Put_x)), 
                                         color = "Breakeven"),
                                         linetype = "dashed") +
  geom_line(data = Put_Payoff_Plot, aes(Put_x, Put_x - NLY_Current_Price, 
                                        color = "Own_Shares"), linetype = "solid") +
  scale_y_continuous(label = scales::dollar) +
  scale_x_continuous(label = scales::dollar) + 
  labs(title = "Buying NLY's $9 Protective Put",
       x = "Stock Price",
       y = "Profit Per Share") +
  scale_color_manual(values = c('Shares'=plot_fg,'Breakeven'=plot_fg,
                                'Breakeven'=plot_fg,"Own_Shares"=plot_fg_alt),
                     labels = c("Protective Put - Profit Per Share (with dividends)",
                                "Owning Shares - Profit Per Share (with dividends)",
                                paste0("Breakeven Price = ", 
                                       currency(NLY_Current_Price + NLY_Put_Cost - 2*NLY_Div)) ),
                     guide = "legend") + 
  guides(color = guide_legend(override.aes = list(linetype = c("solid", "solid", "dashed")))) + 
  options_theme()
```

The black line shows the protective put. It's maximum loss is miniscule. To see why, we'll break down how the difference components of this investment contribute to the maximum loss. We receive `r 2*currency(NLY_Div)` in dividends, and pay `r currency(NLY_Put_Cost)` for the put. We bought the shares for `r currency(NLY_Current_Price)`, and the put has a strike price of `r currency(NLY_Put_K)`. This means the put has `r currency(NLY_Put_K-NLY_Current_Price)` in instrinsic value, which it contributes to the calculation. Putting these together, our maximum loss is `r 2*currency(NLY_Div)` - `r currency(NLY_Put_Cost)` + `r currency(NLY_Put_K-NLY_Current_Price)` = `r currency(2*NLY_Div-NLY_Put_Cost+NLY_Put_K-NLY_Current_Price)`, or just <span class = "fraction" data-top="`r -currency(2*NLY_Div-NLY_Put_Cost+NLY_Put_K-NLY_Current_Price)`" data-bottom="`r currency(NLY_Current_Price)`"></span> = `r -percent( ( 2*NLY_Div-NLY_Put_Cost+NLY_Put_K-NLY_Current_Price ) / NLY_Current_Price )` of our original investment in NLY. The key takeway for an investor is that regardless of what happens to NLY's shares, the most we can lose on January 17, 2020 is `r currency(2*NLY_Div-NLY_Put_Cost+NLY_Put_K-NLY_Current_Price)` per share if NLY pays its two scheduled dividends. 

### Adjusting the Protective Put

I made this trade in my own account because I believed that NLY had a good chance of increasing to \$10 or \$11 by December. If they did, I wanted to sell my \$9.00 January 2020 put and buy a March or April 2020 put with a strike at NLY's new higher price. Since I bought bought NLY at \$8.90, buying a put at a higher strike price would lock in my profits. Buying a put at \$10.00 gives me the right to sell my shares at \$10.00, which locks in a profit of \$1.10 per share. Buying a put at \$11.00 gives me the right to sell my shares at \$11.00, which locks in a profit of \$2.10 per share. 

I bought the protective put on September 24, 2019 and I wanted to roll the put higher in December. If NLY had moved up to \$10 or \$11, selling the old put and buying the new one would cost about \$0.30 per share, less the \$0.25 dividend I get for holding the shares through April. Assuming NLY continued to pay its dividend, adjusting my trade so that I held the longer-dated put would cost around \$0.05 per share. 

In the first leg of the trade I was risking \$0.06 per share to hold NLY until January 20, 2019. If selling the old put and buying the new one costs \$0.05, I'd be risking a total of \$0.11, which is 1.2% of my investment. If NLY had moved to \$10 or \$11 by December, I would guarantee a profit of $1.10 or \$2.10 by April. I bought in September, so that's a 7-month holding period. That's a 7-month return of \$1.10/\$9.06 = 12.14% or \$2.10/\$9.06 = 23.18%. To put that in perspective, those are annual returns of 21.70% and 42.96%. 

If NLY had moved up to \$10, the trade would have a risk-reward of \$1.10-to\$0.11, or 10-to-1. If NLY had moved up to \$11, the trade would have a risk-reward of \$2.10-to-\$0.11, or about 19-to-1. And I'm only risking around 1%, so the upside on this trade if NLY had moved high enough fast enough was enormous. That's what attracted me to the trade. And even if it didn't, the downside was almost nothing. The most I can lose if I decide to exit the trade is 1.2%. 

Alas, it was not to be. NLY had only reached \$9.42 on December 9th, 2019. It was moving higher, but not as fast as I had hoped. The \$9.00 January put was getting close to expiry and I didn't get the move higher that I was looking for. So I knew I had to sell the \$9.00 January 2020 put and buy a $9.00 April 2020 put. On December 9th, I got an opportunity to make that adjusting trade cost-free. The \$9.00 April put fell \$0.05 and the January \$9.00 had not moved. I took the opportunity to *roll* the put's expiry date out to April.

It's now December 19, 2019 and I'm still in this NLY trade, and still only risking \$0.06 per share. I'll want to adjust the \$9.00 April put in March, about 3-4 weeks before its expiry. I still believe that NLY will reach \$10 or \$11, so that I can *roll* to a higher strike put and lock in some profits. For now, the the only thing to do is wait. 


## Managing Your Emotions 

Trading covered calls or protective puts is very different than only trading stocks. If you just own Apple's stock, the best thing for your account is if Apple's shares go up. If they go down, you can just wait until they go back up. As long as you don't need the money for something else and don't sour on Apple as an investment, you can wait as long as it takes for Apple to reach your price target. If Apple goes up, you'll make money and be happy. If Apple goes down, you'll lose money and wish you hadn't made the trade. 


### The Covered Call

The covered call is more complicated. It has two components - the shares you bought, and the call you sold. If you're like most investors, you're more familiar with owning shares than with selling calls. You'll probably view this position as owning shares of Apple with the call tossed on top. And you'll judge the success of writing the call on if it made money compared to just owning Apple's shares. 

Let's look at Apple's $230 covered call that expires in December. Apple is trading at \$222.76, and selling the call nets you \$8.50 per share. You start losing money on the covered call when Apple falls below your breakeven of \$222.76 - \$8.50 = \$214.26. Below \$214.26, you'll always lose \$8.50 less than if you just own shares, because you made \$8.50 from selling a \$230 strike call that expires worthless. Above \$238.50, the call starts eating into gains from owning the shares. You would keep those gains if you had not sold the call. 

The covered call makes money above \$214.26, and makes money compared to just owning shares below \$238.50. Below \$214.26, the covered call helps buffer losses but still loses money. If Apple ends up between \$214.26 and \$238.50 on December 20, 2019, you'll probably view the covered call as successful. Above \$238.50, you'll probably view the call as taking away potential gains you'd get to keep. If Apple is below \$214.26, you might view selling the call as mildly successful because it protected you from greater losses. But you'd proably view the trade as a failure because you still lost money. Are these emotional reactions helpful? I don't think so. 

I think the breakeven (\$214.26 at expiry) and the point where the covered call makes less than owning shares (\$238.50 at expiry) are the key levels you need to know to manage this position. If you sold the covered call and saw Apple rising towards \$238.50 before expiry, you can adjust you position by buying back the call and preserve your upside partcipation in Apple. Everyone has a different view on when to do that - I would only do it when Apple is above \$230 and looking like it was heading above \$238.50. I'd expect the call to cost and extra \$2.00 or so, which I wouldn't mind. Why? Let's go back and look at why I sold the covered call to begin with. 

If I sold the call at \$230, I must have thought Apple didn't have much of a chance of going above \$230 by December 20. If Apple quickly climbs above \$230 after I sell the call, I should buy it back because it means I was wrong. It also means that I've made more money on owning Apple's stock than I thought I would, which should offset the extra \$2.00 I have to pay to buy back the call. If Apple's stock heads lower, it also means I was wrong. I must have thought that the covered call would stay above its breakeven when I sold it, and that I would make money. But I have a lower breakeven than someone who just owns shares, so I have a buffer against these smaller declines. I can use that buffer to close the trade at a profit as it heads towards its breakeven. Someone who just owns shares would have to do so at a loss. 

If I hold through the decline, I'm probably a long-term investor in Apple. In that case, I'm expecting Apple to rebound and I'll keep the \$8.50 I got from selling the call. I'll even sell another one in December, and keep reducing my breakeven on Apple over time. The very unlikely best-case scenario is that I reduce my breakeven to zero so that I'm basically holding the Apple shares for free. I'd still keep writing calls to generate income off of them, which would be amazing if it ever happened. It's more likely that I'd just gradually lower the breakeven over time. 

```{r Apple-205-Covered-Call, out.width = "80%", fig.align = "center", message = FALSE}
Apple_Call_Cost <- 8.50
Apple_Call_K <- 230
x_min <- 190; x_max = 260 
Call_x <- seq(from = x_min, to = x_max, by = 1)

Call_Payoff_Plot <- tibble( "Call_x" = currency(seq(from = x_min, to = x_max, by = 1)),
                            "Covered_Call_y" = currency(Apple_Call_Cost + (Call_x - Apple_Current_Price) - 
                                                  pmax(0, Call_x - Apple_Call_K)) ) 

ggplot() + 
  geom_line(data = Call_Payoff_Plot, aes(Call_x, Covered_Call_y, color = "Shares"), 
            linetype = "solid") +
  geom_line(data = Call_Payoff_Plot, aes(Call_x, rep(0,length(Call_x)), color = "Breakeven"), 
            linetype = "dashed") +
  geom_line(data = Call_Payoff_Plot, aes(x = Apple_Current_Price - Apple_Call_Cost, 
                                         y = seq(from = -4*Apple_Call_Cost, to = x_max-Apple_Call_K, 
                                                 length=length(Call_x)), 
                                         color = "Breakeven"),
                                         linetype = "dashed") +
  geom_line(data = Call_Payoff_Plot, aes(Call_x, Call_x - Apple_Current_Price, color = "Own_Shares"), 
            linetype = "solid") +
  scale_y_continuous(label = scales::dollar) +
  scale_x_continuous(label = scales::dollar) + 
  labs(title = "Selling Apple's $230 Covered Call",
       x = "Stock Price",
       y = "Profit Per Share") +
  scale_color_manual(values = c('Shares'=plot_fg,'Breakeven'=plot_fg,
                                'Breakeven'=plot_fg,"Own_Shares"=plot_fg_alt),
                     labels = c("Covered Call - Profit Per Share",
                                "Owning Shares - Profit Per Share",
                                paste0("Breakeven Price = ", currency(Apple_Current_Price - Apple_Call_Cost))),
                     guide = "legend") + 
  guides(color = guide_legend(override.aes = list(linetype = c("solid", "solid", "dashed")))) + 
  options_theme()
```

Viewing Apple's shares and your covered call as a *single position* makes reasoning about the trade much easier. Selling the \$230 strike call skews the risk-reward proposition of the entire trade. The payoff diagram that compares the covered call to just owning shares summarizes how in one quick picture. All you have to do to manage the trade is remember why you entered it - to collect some income and reduce your volatility - and look at the payoff diagram. It clearly shows that the covered call outperforms just owning shares below \$238.50. So you should prefer the covered call to just owning shares below \$238.50, and close the call if Apple starts moving towards \$238.50 before expiry. It also shows that the covered call sustains losses below \$214.26. Depending on your tolerance for these declines, you might want to close the covered call (sell the shares and buy back the call) as Apple moves towards its breakeven. Doing so while Apple is sufficiently above \$214.26 lets you exit the position at a profit. 

There are many ways to risk-mange a covered call. This discussion only highlights two key levels on the payoff diagram - buying back the call as Apple moves towards \$238.50 to avoid losing upside partcipation in Apple, and closing the entire position (the shares and the call) as Apple moves towards \$214.26. Apple was trading at \$222.76 when you sold the call, so you have roughly \$8.50 to the downside and \$15.74 to the upside before having to make an adjusting trade. We'll discuss the payoff diagrams of options before their expiry date in detail later, but the key thing to know is that options don't respond as much to price movements in their stocks as they further away from their expiry date. That means your breakeven will likely be higher than \$214.26 before expiry, as the call will not have lost all its value as Apple's shares go down. The \$238.50 will be higher, as the call will not have gone up as fast as Apple's shares head higher.

<!-- Turn this into a diagram: the window is about \$25 because it uses the \$8.50 you get from selling the call, as well as the *deductible* of \$230 - \$222.76 = \$7.24 before the upside protection from the call kicks in. -->
In addition to providing income and reducing your volatility, the covered call gives you more *flexibility* than just owning shares. This flexibility lets you adjust the trade in ways that who just owns shares can't. By viewing the covered call as a *single* position with different risk-rewards characteristics than just owning shares and exploiting this *fliexibility* lets you take your emotions out of the trade. For my risk preferences, I would close the entire position before it reaches its breakeven and buy back the call as Apple approaches \$238.50. If you have different risk preferences - if you're okay with losing a few percent on the trade or don't mind losing the upside in Apple - then you'll have different levels. But you'll still have more flexibility to manage the position than someone who just owns shares. And you'll be able to keep your emotions from interfering with profitably executing the covered call. 


### The Protective Put 

People tend to think about the protective put in the same way, as two separate positions. The primary position is the stock, and the put is an add-on that protects them from potential downside risks. But the put costs money, so if the stock doesn't fall below its strike price, you might think you wasted money buying the put. Because it will expire worthless and not do anything useful. 

Just like with the covered call, I'm going to argue that the protective put should be viewed as a *single* position with different risk-reward characteristics than just owning shares. Using the payoff diagram will guide us towards the levels we need to make adjusting trades to preserve the risk-reward characteristics that can make the protective put such an attractive trade. This time, I'll demonstrate with the NLY trade. 

```{r NLY-Protective-Put-Manage, out.width = "80%", fig.align = "center", message = FALSE}
x_min <- 7; x_max <- 11
Put_x <- seq(from = x_min, to = x_max, by = 0.1)
Put_Payoff_Plot <- tibble( "Put_x" = currency(seq(from = x_min, to = x_max, by = 1)),
                            "Protective_Put_y" = currency(-NLY_Put_Cost + 2*NLY_Div +
                                                            ( Put_x - NLY_Current_Price ) + 
                                                            pmax(0, NLY_Put_K - Put_x)) ) 

ggplot() + 
  geom_line(data = Put_Payoff_Plot, aes(Put_x, Protective_Put_y, 
                                        color = "Shares"), linetype = "solid") +
  geom_line(data = Put_Payoff_Plot, aes(Put_x, rep(0,length(Put_x)), 
                                        color = "Breakeven"), linetype = "dashed") +
  geom_line(data = Put_Payoff_Plot, aes(x = NLY_Current_Price + NLY_Put_Cost - 2*NLY_Div, 
                                         y = seq(from = -4*NLY_Put_Cost, 
                                                 to = 3*(x_max-NLY_Put_K), 
                                                 length=length(Put_x)), 
                                         color = "Breakeven"),
                                         linetype = "dashed") +
  geom_line(data = Put_Payoff_Plot, aes(Put_x, Put_x - NLY_Current_Price, 
                                        color = "Own_Shares"), linetype = "solid") +
  scale_y_continuous(label = scales::dollar) +
  scale_x_continuous(label = scales::dollar) + 
  labs(title = "Buying NLY's $9 Protective Put",
       x = "Stock Price",
       y = "Profit Per Share") +
  scale_color_manual(values = c('Shares'=plot_fg,'Breakeven'=plot_fg,
                                'Breakeven'=plot_fg,"Own_Shares"=plot_fg_alt),
                     labels = c("Protective Put - Profit Per Share (with dividends)",
                                "Owning Shares - Profit Per Share (with dividends)",
                                paste0("Breakeven Price = ", 
                                       currency(NLY_Current_Price + NLY_Put_Cost - 2*NLY_Div)) ),
                     guide = "legend") + 
  guides(color = guide_legend(override.aes = list(linetype = c("solid", "solid", "dashed")))) + 
  options_theme()
```

Assuming NLY continues to pay its dividend, I entered the NLY trade only risking \$0.06 per share, which is 0.67% of my investment. The most I can lose on the expiry date of my put is 0.67%. If NLY reaches \$10, the trade will have a risk reward of 10-to-1 and if NLY reaches \$11, the trade will have a risk-reward of about 19-to-1. This is *dramatically* different than if I just owned shares of NLY. In that case, I could lose everything I invested into NLY. 

In this example, the protective put changes the risk-reward characteristics from potentially losing everything to *knowing* that you can only lose less than 1% over the lifetime of the trade. This example is especially convincing because NLY is paying a 10% dividend, which essentially pays for the put. I'm using the put to remove the downside risk of owning NLY shares, and then using the dividend to pay for almost all of the put. 

Even when the put is more expensive, it limits your losses to 0.67% (as long as NLY pays their dividend). Psychologically, this removes the stress that comes with investing in an otherwise risky stock. It completely changes how you think about being invested in NLY. Instead of worrying about how much money you could lose if NLY goes down, you have an essentially stress-free position in a stock with lots of potential upside. All you have to do is sit and wait to see if NLY goes higher. 

Just like the covered call, viewing the protective put as a *single* position gives you more *flexibility*. The flexibility is simpler to see with the protective put - if NLY goes down, you don't lose money. You can close the position and wait for NLY to stop falling, and then buy back into it when it does. Or just buy back into it right away, at the lower price. You'll still be ahead of someone who just owned the stock. 

I think most people are more concerned with what could happen if they lose lots of money in the stock market than they are with missing out on thes few percent they could get by selling a call. I hope this NLY example demonstrates how amazing protective puts can be. They allow you to define the maximum loss you are willing to tolerate on an investment, which has significant practical and psychological benefits. Protective puts also allow your to construct trades with superior risk-reward characteristics compared to just owning stocks. If you find your investment portfolio is causing you too much stress, you might want to discuss if protective puts can help take some of the stress out of investing with an investment advisor. 


## Synthetic Positions {#Synthetics}

Understanding synthetics is one of the most important things you can do to understand the risks of an option position. You can find what the synthetic of any position is on Google, so I'm not going to create an encyclopedia of them here. But I'm going to go through a few examples of using synthetics to better understand the risks of an options position, so that you can see how synthetics can help you understand these risks. 

### The Combination 

Buying a stock consists of buying 100 shares. But it has the *same* payoff diagram as buying a call and selling a put, where they have the same strike price and expiry date. First, let's combine buying a call and selling a put to see what that payoff diagram looks like. 

<!-- Insert two payoff diagrams (at expiry) on the same graph: a long call and short put (same strike) --> 

This position is called a *combination* or *combo*. Let's compare the payoff diagram of owning shares to owning the combination. They are, in fact, the same. 

<!-- Insert two payoff diagrams (at expiry) side-by-side: a combination and long stock --> 

There are three pieces at work here: the shares, the call, and the put. And in fact, any position taken using some of those three pieces can be *replicated* using the other pieces. If you replicate a position using options, you've created a *synthetic* of the original position. For example, buying a call and selling a put is a *synthetic* of buying shares. 

Bu even though the payoff diagrams of owning stock and owning a combination look the same, the positions are *not* identical. The most important difference is that the options expire, while the shares do not. I've shown the payoff of the combination at expiry, when it matches the payoff of the shares. It does *not* match the payoff of the shares before expiry or after expiry. So in this case, the payoff diagrams *only* match on the expiry date of the combination.

The combination says that owning shares is like owning a call and selling a put. How can this help us understand the risks of owning shares? 


### The Covered Call

What about the covered call? The covered call consists of buying 100 shares and selling a call. But it has the *same* payoff diagram as selling a put (with the same strike as the call). Take a look: 

<!-- Insert two payoff diagrams side-by-side: The $230 Apple covered call and the $230 Apple short put. -->

That means the risks (and rewards) of the covered call are the same as selling a put. Let's forget everything I said about the covered call being a viable income-production strategy that mildly reduces volatility, and start over by selling a put at the same strike. That's the December 20, 2019 put on Apple with a \$230 strike, which costs \$15.30. If I sell this \$230 strike put, my payoff diagram looks like this:

<!-- Insert payoff diagram: short put of Apple's $230 strike (cost: $15.30) --> 

If I sell this put and Apple's shares are below \$230 on December 20, the person I sold them to can make money by exercising the put and forcing me to buy them for \$230. Just like selling a *naked* call could force me into a mad scramble for cash, selling a *naked* put can do the same thing. If I just sell the put and it gets exercised, I have to come up with the cash to buy the shares. If I don't have the cash, this can be completely disasterous. 

There are two ways I can protect myself. I can write a covered put, where I short the stock and sell the put. The payoff diagram of the covered put makes this clear: if Apple's stock goes down, any losses I sustain from having sold the put will be covered by the gains I make by being short the stock. In other words, being short the stock *covers* losses I would otherwise take by selling the put. 

<!-- Insert diagram: covered put -->

The other way I could protect myself is by writing a cash-secured put. This is where I keep enough cash in my account to buy the shares if I needed to. For this \$230 strike put, I would need to leave \$230 * 100 = \$23,000 in my account to cover the worst-case scenario of the put being exercised at \$229.99. If Apple's shares fall below \$230, the put can be exercised early so I should have that cash in my account the entire time. Cash secured puts may have a place in other people's investing strategies, but locking up that much cash until the put expires doesn't appeal to me very much. 

Moving back to the covered call: I said the covered call has a number of desirable properties. It reduces the volatility of your investments, and it is a reasonable income-enhancing strategy. You can collect income by writing calls against investments you already own, which has a certain appeal. This makes the covered call sound like a safe strategy, but it's payoff diagram is the *same* as a naked put. Is the naked put (or even the cash secured put) safe? 



<!--
## The Straddle
The straddle is an options strategy where an investor buys a call and a put, where both have same expiry and same strike price. Let's begin by looking at the payoff diagram of a \$220 strike straddle on Apple's shares. This time, we bought the shares at $219.84 on September 26, 2019. The put costs \$11.10, the call costs \$11.25, and they both expire on December 20, 2019. 

```{r Apple-205-Strike-Straddle, echo = FALSE, out.width = "80%", fig.align = "center", message = FALSE}
Call_cost <- 11.25
Put_cost <- 11.10
Call_K <- 220; Put_K <- 220
Call_x <- seq(from = 160, to = 260, by = 0.01)
Put_x <- seq(from = 160, to = 260, by = 0.01)

plot(x = Call_x, 
     y = -Call_cost - Put_cost + pmax(0, Call_x - Call_K) + pmax(0, Put_K - Put_x), 
     type = 'l', lwd = 3, ylim = c(-75, 75), col = "black",
     xlab = "Apple's Stock Price", ylab = "Profit Per Share",
     main = "Payoff of a Straddle on Expiry")
lines(x = Call_x, y = Call_x - 219.84, col = "orange", lwd = 3)

abline(v = 220 + Call_cost + Put_cost, lty = 2, lwd = 1, col = "black")
abline(v = 220 - Call_cost - Put_cost, lty = 2, lwd = 1, col = "black")
abline(h = 0, lty = 2, lwd = 1)

# Intersection between owning shares and owning the straddle - dollars
# Intersect_dollars <- ( -Call_cost - Put_cost + pmax(0, Call_x - Call_K) + pmax(0, Put_K - Put_x) ) - (Call_x - 219.84)
# Call_x[which( Intersect_dollars == Intersect_dollars[which.min(abs(Intersect_dollars))])] # Call_x[3251] = 208.75

legend("topleft", c("Straddle - Profit Per Share",
                    paste0("Breakeven Price (Down) = $", 
                           219.84 - Call_cost - Put_cost),
                    paste0("Breakeven Price (Up) = $", 
                           219.84 + Call_cost + Put_cost),
                    "Owning Stock - Profit Per Share"),
       lty = c("solid", "dotdash", "dotdash", "solid"), 
       lwd = c(2, 2, 2, 2), cex = 0.75,
       col=c("black", "black", "black", "orange"), bty = "n") 
```

This is the most complicated payoff diagram we've seen so far. The first thing to notice is that we bought Apple's shares at \$219.84 and bought puts and calls with a \$220 strike. This was the closest strike to our purchase price, which is what we want to purchase if we want to profit if there is a large move in Apple's stock. The major benefit of the straddle is that it gives us the flexibility to not have to bet on the direction of that move. The call will profit if the stock goes up, and the put will profit if the stock goes down. The major downside of the straddle is that we have to buy two options (a put and call) instead of one (a put or a call). That means that we must recover the cost of both options before we become profitable. If the stock does not have a large move in either direction, we will lose money. 

Since we bought a put and a call, we can profit if the stock goes up a lot or down a lot. Each direction has its own breakeven price. We spent \$11.25 on the call and \$11.10 on the put, for a total of \$11.25 + \$11.10 = \$22.35. We bought Apple's shares for \$219.84. For the put to breakeven, the shares must fall by \$22.35 to \$219.84 - \$22.35 = \$197.49. For the call to breakeven, the shares must climb by \$22.35 to \$219.84 + \$22.35 = \$242.19. These are the two breakeven prices shown on the payoff diagram. 

The most we can lose is \$29.08, since we paid that much for the two options. This maximum loss happens when Apple's shares are at \$220, since neither option makes money when they expire at \$220 Each dollar we move in either direction recovers a dollar of the cost of the straddle, and every dollar we move beyond either breakeven price is a dollar of profit. Just owning shares (the orange line) is more profitable than owning a straddle once Apple's shares are above 208.74. Below \$208.75, the shares continue to lose value and the put gains value, making the stradde the better investment. Above \$208.75, the shareholder performs better on account of not having to pay for two options. 

The payoff diagram makes it easy to see the intuition behind the straddle. At expiry, it loses money between \$197.49 and \$242.19. These are roughly 10\% below and above the \$219.84 we bought Apple's shares for. This gives us the intuition that Apple's shares must move around 10\% (in either direction) for this straddle to be profitable. We'll make that intuition firmer when we discuss how to construct options trades exploit their prices movements before their expiry date, but our goal for now is getting the big-picture ideas behind each options strategy without getting bogged down in details. For the straddle, the big picture idea is that the straddle lets us bet on a big price move but not bet on its direction. 


## Replicating Portfolios
You might have noticed that the payoff diagram of a protective put looks pretty close to the the payoff diagram of a call. Let's take another look when the put and call both have a \$205 strike and expire on December 20, 2019. Like some of our previous examples, we'll suppose that we bought Apple's shares at \$222.76 on September 11, 2019. The call is shown in blue, the protective put is red, and just owning shares is orange. The red and blue lines largely overlap, so they may be difficult to see. 

```{r Replicate-Expiry-Payoffs, echo = FALSE, fig.height = 4, fig.width = 6, fig.align = "center"}
Put_cost <- 5.45
Call_cost <- 23.63
Put_K <- 205; Call_K <- 205
Put_x <- seq(from = 180, to = 260, by = 1)
Call_x <- seq(from = 180, to = 260, by = 1)

plot(x = Put_x, y = -Put_cost + ( Put_x - 222.76 ) + pmax(0, Put_K - Put_x), 
     type = 'l', lwd = 3, ylim = c(-30, 30), 
     col = alpha(rgb(255,0,0,maxColorValue=255),0.5),
     xlab = "Apple's Stock Price", ylab = "Profit Per Share",
     main = "A Replicating Portfolio - On The Expiry Date")
lines(x = Put_x, y = Put_x - 222.76, col = "orange", lwd = 3)

abline(v = 222.76 + Put_cost, lty = 2, lwd = 1, col = "red")
abline(h = 0, lty = 2, lwd = 1, col = "red")

lines(x = Put_x, y = - Call_cost + pmax( 0, Call_x - 205 ), lwd = 3, 
      col = alpha(rgb(0,0,255,maxColorValue=255), 0.5) )
abline(v = 205 + Call_cost, lty = 2, lwd = 1, col = "blue")
abline(h = 0, lty = 2, lwd = 1, col = "blue")

legend("topleft", c("Protective Put - Profit Per Share",
                    paste0("Protective Put - Breakeven Price = $", 222.76 + Put_cost),
                    "Call - Profit Per Share", 
                    paste0("Call - Breakeven Price = $", 205 + Call_cost),
                    "Owning Stock - Profit Per Share"),
       lty = c("solid", "dotdash", "solid", "dotdash", "solid"), 
       lwd = c(2, 2, 2), cex = 0.75,
       col=c("red", "red", "blue", "blue", "orange"), bty = "n") 
```

The first thing to notice is that the call has a slightly higher breakeven than the protective put - \$228.63 versus \$228.21. This makes sense, because we have to spend \$228.63 to buy the protective put, but only \$23.63 to buy the call. The difference of \$228.63 - \$23.63 = \$205 is sitting in the call owner's bank account earning interest. Savings accounts are earning around 1\% per year, whcih is 0.249\% every 3 months. We bought the options on September 11, 2019 and they expire on December 20, 2019, in about 3 months' time. The call owner will make an additional 3 months of interest from the cash in their savings account, which comes out to \$205 * 0.249\% = $0.51. This lowers the breakeven of the call by \$0.51, from \$228.63 to \$228.63 - \$0.51 = \$228.12. This is within pennies of the \$228.21 breakeven of the protective put. 
-->
