<!-- Each chapter is set to compile separately - include "global" set-up -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

CRAN_repo = getOption("repos")
CRAN_repo["CRAN"] = "http://cran.us.r-project.org"
options(repos = CRAN_repo)

if ( !require("here") )          { install.packages("here") };            require(here)
if ( !require("tidyverse") )     { install.packages("tidyverse") };       require(tidyverse)
if ( !require("ggtext") )        { install.packages("ggtext") };          require(ggtext)
if ( !require("pBrackets") )     { install.packages("pBrackets") };       require(pBrackets)
if ( !require("htmlwidgets") )   { install.packages("htmlwidgets") };     require(htmlwidgets)
if ( !require("plotly") )        { install.packages("plotly") };          require(plotly)
if ( !require("formattable") )   { install.packages("formattable") };     require(formattable)
if ( !require("kableExtra") )    { install.packages("kableExtra") };      require(kableExtra)
if ( !require("rtsdata") )       { install.packages("rtsdata") };         require(rtsdata)
if ( !require("PerformanceAnalytics") )     { install.packages("PerformanceAnalytics") };   require(PerformanceAnalytics)
if ( !require("anytime") )       { install.packages("anytime") };         require(anytime)
if ( !require("gridExtra") )     { install.packages("gridExtra")};        require(gridExtra)


options(kableExtra.html.bsTable = TRUE)
options(kableExtra.auto_format = FALSE)
```

```{r Set-Global-Chapter-Variables}
plot_bg <- "#FFFFFF" # "#191919"
plot_fg <- "#000000" # #929292"
plot_fg_alt <- "#969696" # ??
vline_color <- "#D2B48C"
  
save_tables_as_pictures <- FALSE
load_table_pictures <- TRUE
out_width_plot <- "80%"
out_width_table <- "90%"

grid::grid.locator(unit="npc") 
output_epub = knitr::opts_knit$get("rmarkdown.pandoc.to") == "epub3"
source("Plotting_Functions.R")
```
 

```{r Table-Helper-Functions}
at_least_number <- function(x,number) {
  ifelse( typeof(x) == "character", 
          return( as.numeric(gsub('\\(','-',gsub('[)$,]', '', x))) >= number ),
          return( x >= number) )
}

remove_negative_zeros <- function(x) { (x[which(-1e-4 < x & x < 1e-4)] = 0); return(x) }
```

# The Straddle Revisted 

We broke down the straddle in the last section. A standard straddle buys at-the-money calls and puts on an index, like the S&P 500. The thinking is that the call makes you money if the S&P goes up, and the put makes you money if the S&P goes down. It has to move *enough* so that you cover the cost of entering the straddle, but you're profitable on (large enough) upside upside and downside moves in the market. The problem is that you burn time value -- you've bought puts and calls on the S&P, and they both lose value while you sit around and wait for the S&P to move. Ideally, the S&P will move enough to pay for the time value you lose, but that may not happen. 

One strategy that I've used with some success in a dividend-focused portfolio is to allocate 15%-20% to 2-3 high beta positions, and allocate 1% each to near-the-money calls on those positions. This lets me keep 80% of the portfolio invested in dividend-paying stocks, while participating in much more of the potential upside in these high beta positions. It also limits my losses to the cost of the calls. I've found this strategy gives my portfolio compelling upside potential, while letting me maintain its focus on steady dividend growth. I'd like to be able to allocate 10% of the portfolio to calls on high-beta stocks, but those calls could get wiped out in a market correction. I think that would compromise the slow-and-steady approach of the dividend portfolio, so I have to to solve that problem. 

The straddle hints at a solution -- it uses puts to hedge the downside. It's big drawback is that it probably costs money to carry, which makes it undesirable as a long-term allocation. If I can increase my chances of carrying it at a profit, the straddle becomes more appealing. One way to do that is to notice that the S&P 500 doesn't move very much if one of its stocks does. Apple is the biggest weight in the S&P in winter 2021, making up around 6% of the index. If Apple goes up 10%, it only takes the S&P up 0.6%. That's not enough to for the call to make much money. If I had bought calls on Apple instead, that 10% move would have made me a lot of money. That's exactly what my goal is - I want to calls to capture more of upside in individual names, while still keeping the put so that I can protect my portfolio against against market declines.

That's the modified straddle - buy puts on an index and buy calls on a basket of stocks in the index. I've mentioned the S&P 500, but it could just as easily be the NASDAQ, the Russell, or any other index. It just depends on what stocks you want to buy calls on. The risk-reward of the modified straddle differs quite a bit from the standard straddle. Let's look at straddles and modified straddles that I've run in my own accounts, to see how their risk-reward profiles and how risk-managing them is very different. 

*Straddles*

*Modified Straddles*


