<!-- Each chapter is set to compile separately - include "global" set-up -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Source: https://davidgohel.github.io/officedown/articles/captions.html
knitr::opts_chunk$set(
  fig.cap.style = "Image Caption",
  fig.cap.pre = "", fig.cap.sep = "")

knitr::opts_chunk$set(
  tab.cap.style = "Table Caption",
  tab.cap.pre = "", tab.cap.sep = "")

CRAN_repo = getOption("repos")
CRAN_repo["CRAN"] = "http://cran.us.r-project.org"
options(repos = CRAN_repo)

if ( !require("here") )          { install.packages("here") };            require(here)
if ( !require("tidyverse") )     { install.packages("tidyverse") };       require(tidyverse)
if ( !require("ggtext") )        { install.packages("ggtext") };          require(ggtext)
if ( !require("pBrackets") )     { install.packages("pBrackets") };       require(pBrackets)
if ( !require("htmlwidgets") )   { install.packages("htmlwidgets") };     require(htmlwidgets)
if ( !require("plotly") )        { install.packages("plotly") };          require(plotly)
if ( !require("formattable") )   { install.packages("formattable") };     require(formattable)
if ( !require("kableExtra") )    { install.packages("kableExtra") };      require(kableExtra)
if ( !require("grid") )          { install.packages("grid") };            require(grid)
if ( !require("magick") )        { install.packages("magick") };          require(magick)
if ( !require("gridExtra") )     { install.packages("gridExtra") };       require(gridExtra)
if ( !require("derivmkts") )     { install.packages("derivmkts") };       require(derivmkts)

options(kableExtra.html.bsTable = TRUE)
options(kableExtra.auto_format = FALSE)

out_width_plot <- "97.5%"
out_width_table <- "97.5%"
grid::grid.locator(unit="npc") 

plot_bg <- "#FFFFFF"
plot_fg <- "#000000" 
plot_fg_alt <- "#969696" 

source("Plotting_Functions.R")
```

```{r Table-Helper-Functions}
at_least_number <- function(x,number) {
  ifelse( typeof(x) == "character", 
          return( as.numeric(gsub('\\(','-',gsub('[)$,]', '', x))) >= number ),
          return( x >= number) )
}

remove_negative_zeros <- function(x) { (x[which(-1e-4 < x & x < 1e-4)] = 0); return(x) }
```

# LEAPS

## Rho

As the price of XYZ's shares change, the price of its options do. *Delta* and *gamma* are two tools we can use to understand these changes. When we start constructing option trades, thinking through the *delta* and *gamma* exposure that each position gives us will be an important risk management tool. There are 3 more main factors that we have to risk manage in an options position -- *rho*, *vega*, and *theta*.  We'll cover *rho* here, and *vega* and *theta* in the next section. 

Rho measures how much the price of an option changes because of changes in interest rates. Why do interest rates influence the price of options? The idea comes from considering how to construct the same trade, but without using options. For a call...


## Why Use LEAPs?

