<!-- Each chapter is set to compile separately - include "global" set-up -->
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

if ( !require("tidyverse") )     { install.packages("tidyverse") };       require(tidyverse)
if ( !require("htmlwidgets") )   { install.packages("htmlwidgets") };     require(htmlwidgets)
if ( !require("plotly") )        { install.packages("plotly") };          require(plotly)
if ( !require("grDevices") )     { install.packages("grDevices") };       require(grDevices)
if ( !require("formattable") )   { install.packages("formattable") };     require(formattable)
if ( !require("kableExtra") )    { install.packages("kableExtra") };      require(kableExtra)
if ( !require("derivmkts") )     { install.packages("derivmkts") };       require(derivmkts)

options(kableExtra.html.bsTable = T)
options(kableExtra.auto_format = FALSE)
```

```{r Set-Global-Chapter-Variables, echo = FALSE}
plot_bg <- "#FFFFFF" # "#191919"
plot_fg <- "#000000" # "#929292"
plot_fg_alt <- "#969696" # ??
label_font_size <- 7.5

options_theme <- function() {
  theme_classic() + 
    theme(plot.title = element_text(hjust = 0.5, face="bold", 
                                    margin = margin(10, 0, 10, 0),
                                    colour = plot_fg),
          legend.justification=c(0,1), 
          legend.position=c(0,1), 
          legend.title = element_blank(),
          legend.key = element_rect(fill = plot_bg, color = plot_bg),
          legend.background = element_rect(fill = plot_bg, color = plot_bg),
          legend.box.background = element_rect(fill = plot_bg, color = plot_bg),
          legend.text = element_text(colour = plot_fg),
          text = element_text(size=16), 
          panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          panel.background = element_rect(fill = plot_bg, color = plot_bg),
          plot.background = element_rect(fill = plot_bg, color = plot_bg),
          axis.line = element_line(colour = plot_bg),
          axis.title.x = element_text(margin = unit(c(4, 0, 0, 0), "mm"), 
                                      colour = "black"),
          axis.title.y = element_text(margin = unit(c(0, 4, 0, 0), "mm"), 
                                      angle = 90, colour = "black"),
          axis.text.x = element_text(colour = "black"), 
          axis.text.y = element_text(colour = "black") ) 
}

no_axis_ticks <- function() {
  theme( axis.title.x = element_blank(), 
         axis.text.x = element_blank(), 
         axis.ticks.x = element_blank(),
         axis.title.y = element_blank(), 
         axis.text.y = element_blank(), 
         axis.ticks.y = element_blank() )
}
```

# Using the Greeks {#Using-the-Greeks}

The *greeks* are the risk factors of an options position. If you ever take a position in options, they are what will cause you to gain or lose money. So you should understand the exposure to the *greeks* that any options trade you make has. We've covered 5 *greeks* so far. Here's a quick recap of all 5:  

  1. *Delta* - how much the option's price changes when the stock goes up by \$1. *Delta* is positive for calls and negative for puts. 
  2. *Gamma* - how much *delta* changes with the next \$1 move in the stock. *Delta* will increase if the \$1 move causes the option to gain intrinsic value and decrease if it causes the option to lose intrinsic value. 
  3. *Vega* - how much the option's value changes with a 1 point change in implied volatility. 
  4. *Theta* - how much *time value* the option will lose the next calender day. 
  5. *Rho* - how much the option's value changes with a 1 point change in the risk-free interest rate. 

There are more *greeks*. I've only ever used three other *greeks* when trying to construct trades. I've never actually executed trades that used these three *greeks*, since the strategies that I know that use them are quite complex. In case you're wondering, those three *greeks* are: 

  1. *Vanna* - how much the *delta* of an option changes with a 1 point change in implied volatility. 
  2. *Volga* - how much the *vega* of an option changes with a 1 point change in implied volatility. 
  3. *Speed* - how much *gamma* changes with a \$1 move in the stock. Also called the *gamma of gamma*.  

It's almost hard to fathom that we could need even more *greeks*, but I'll point out where these three could come in handy while walking through options strategies that use the 5 *greeks* that I've already covered. 

Once you know how various options give you different exposures to those 5 *greeks*, you can construct options trades that *hedge* out one or more of the *greeks*. For example, if I didn't want have risk exposure to *delta*, I could construct an options trade that is *delta* hedged. Similarly, I could construct options trades that are *vega* hedged, *gamma* hedged, *delta* hedged and *gamma* hedged, and so on. I'll walk through examples of options strategies that hedge out various *greeks*, and make sure to discuss what types of market environments they perform well in and how some things to keep in mind while managing the trade. 

But first, let's summarize how the *greeks* affect calls and puts. These lists are good references to look back at if you're 

## Recapping the Greeks {#Recap-the-greeks}

If you buy a call, you are: 

  1. Long *delta* - if the stock goes up by \$1, the price of the call goes up by *delta*.  
  2. Long *gamma* - if the stock goes up by \$1, the *delta* of the call goes up by *gamma*.
  3. Long *vega* - if the implied volatlity of the call goes up by 1, the price of the call goes up by *vega*. 
  4. Short *theta* - if you own the call, you pay time decay. 
  5. *Rho* - if the , the price of the call goes up by  . 

If you buy a put, you are:  

  1. Short *delta* - if the stock goes up by \$1, the price of the put goes down by *delta*.  
  2. Long *gamma* - if the stock goes up by $1, the *delta* of the call goes down by *gamma*. 
  3. Long *vega* - if the implied volatlity of the call goes up by 1, the price of the call goes up by *vega*. 
  4. Short *theta* - if you own the put, you pay time decay. 
  5. *Rho* - if the , the price of the put goes up by . 

If you sell a call, you are:

  1. Short *delta* - if the stock goes up by \$1, the call you sold goes up. For you, that leads to a loss. 
  2. Long *gamma* - . 
  3. Short *vega* - . 
  4. Long *theta* - if you sell the call, you benefit from time decay. 
  5. *Rho* - . 

If you sell a put, you are: 

  1. Long *delta* - if the stock goes up by \$1, the put you sold goes down. For you, that leads to a gain. 
  2. Long *gamma* - . 
  3. Short *vega* - . 
  4. Long *theta* - if you sell the put, you benefit from time decay. 
  5. *Rho* - . 


## The Straddle 

The straddle is a *delta* hedged strategy. It consists of buying 1 put and buying 1 call, both with the same expiry date and are at-the-money. It's payoff diagram looks like this: 

<!-- Insert payoff diagram: The straddle --> 

Let's take a look at a straddle trade I've done in the past:


## The Collar 

The collar is a ... . It consists of buying 100 shares, buying 1 put, and selling 1 one call. It' s


## The Calender Spread




## The Conversion / Reversal

The conversion is a ..., while the reversal is ... . 